Haploid Selection, Sex Ratio Bias, and Transitions Between Sex-Determination Systems

### 2018-05-01 LaTeX/response.tex

\noindent\subsection{}
Likewise, the results yield three major insights:\\
(1) SD transitions can, counterintuitively, potentially INCREASE sex ratio bias\\
(2) Tight linkage and sexual antagonism are not required for transitions to occur\\
(3) Under haploid selection, conditions for the spread of new sex-determining alleles are extremely permissive.\\
Again, all of these points are there, but they could be better drawn out than they currently are. The results, as written, are hard-going, and I expect will be particularly challenging for readers that are not heavily immersed in evolutionary theory.

One potential way to help readers -- particularly those interested in what the paper has to say, but are not mathematically inclined -- is to conspicuously highlight each important result and briefly place it into context of earlier theory. But at the same time, the authors have done a great job of drawing out general mathematical results from a relatively complicated model -- their mathematical insights ARE the results and I am not suggesting that they be minimized or shunted into appendices.

One tactic for including analytical details which specialists will want to see, while allowing less mathematically inclined readers to easily find each main result, is to place short and accessible summaries at the close of each section of the Results. Such a tactic has been put to good use in a hugely influential paper by Allen Orr (1998 Evolution), which I expect the authors will be familiar with. Orr presents the major mathematical results of his analysis alongside prominently labelled punchline paragraphs that summarize each of his main results in a concise and easily understandable manner. If the authors choose to take this approach, please be clear to place the take-home messages in context: clearly state whether the result confirms previous theory, or represents a new result or contradiction of an earlier result.

\noindent\paragraph{Response}
\textcolor{blue}{
%We now give 3 main conclusions at the end of the introduction (L69) which echo those mentioned by the reviewer (looser linkage can evolve, equal selection to increase or decrease the sex ratio, ESD does not always invade biased sex ratio; all of which suggest haploid selection makes conditions for transitions more permissive).
Following this advice, we now highlight our main conclusions in the results section in the style of Orr 1998:
\begin{enumerate}
\item[(1)] ancestrally-tight sex-linkage allows trans-GSD transitions but not cis-GSD transitions (L228-230)
\item[(2A)] with ancestrally-tight sex-linkage, trans-GSD transitions (which weaken sex-linkage and can lead to polymorphic sex determination systems) can be driven by selection to better specialize on the ancestrally-non-heterogametic sex (L282-286)
\item[(2B)] with ancestrally-tight sex-linkage, haploid selection can spur trans-GSD transitions (which weaken sex-linkage and equalize the sex-ratio) even when selection is sexually-antagonistic in diploids  (L288-293)
\item[(3A)] when selection is weak relative to ancestral sex-linkage, cis- or trans-GSD transitions can occur when the new sex-determining allele increases linkage with selected loci (L324-326)
\item[(3B)] when selection is weak relative to ancestral sex-linkage, haploid selection allows trans-GSD transitions to occur even when the new sex-determining allele decreases linkage with the selected locus (L327-330)
\item[(4A)] when selection is weak relative to ancestral sex-linkage, sex-ratio selection is exactly equal in magnitude to the selection imposed by associations with selected alleles (and thus transitions can increase or decrease the sex-ratio bias) (L350-353)
\item[(4B)] transitions from genetic to environmental sex-determination are not straightforwardly predicted by selection to balance the zygotic sex ratio when haploid selection is present (L411-413)
%\vspace{0.25cm}\\
\end{enumerate}
%As (1A), (1B), and (2) have some antecedents in the literature, we discuss these immediately below each conclusion. 
%We also discuss, immediately below each conclusion, its novelty.
%These are all new results, although (1A) was shown for the special case of $r=0$, $R=1/2$, and ploidally-antagonistic selection in Kozielska et al 2010, which we now mention immediately below (1A).
\textcolor{red}{what is new and what is confirmatory? [see red notes in main text in conclusions]} 
}

@@

\section{Reviewer 3}

This work (henceforth [SOO]) explores the role of both diploid and haploid selection on the evolution of new sex-chromosomes and or transitions between sex-determination systems. For that purpose, the authors develop a population genetics model of three loci with two alleles in each locus. The first locus X is the ancestral sex-determining locus, the second locus A is the locus under diploid and haploid selection. The third locus M is the novel sex-determining locus.

\noindent\subsection{}
I find the article is not clearly written and could benefit from improving the presentation of the main results. 
%
%\noindent\paragraph{Response}
%\textcolor{blue}{We hope the improvements discussed above (e.g., Response 2.1 and 2.2) and below have increased the clarity of our writing, especially regarding the main results.}
%
%\noindent\subsection{}
I also find the article claims that: (a) the modeling work is novel and (b) the findings are unexpected to be only partially justified.
%
%\noindent\paragraph{Response}
%\textcolor{blue}{As described at length below, we now give more weight to two previous papers (Kozielska et al 2010 and Ubeda et al 2015) and more precisely indicate which of our results are new and unexpected.}
%
%\noindent\subsubsection{}
The works of Kozielska 2014 and Ubeda 2015 model the combined effects of both haploid and diploid selection on the evolution of new sex-chromosomes. The work of [SOO] is thus not novel in considering both haploid and diploid selection on the evolution of new sex-chromosomes. It does generalize previous work by considering sex-specific viability selection and get some nice analytical results.

\noindent\paragraph{Response}
%\textcolor{red}{The abstract and summary still seem to be a little over-selling...}
\textcolor{blue}{
We have performed an extensive re-arrangement to improve the clarity of our writing, especially regarding the main results. 
As described in Response 2.1 (point 6) and below, we now provide a prominent and longer discussion of Kozielska et al 2010 and Ubeda et al 2015 (L57-69) and more precisely indicate which of our model features and results are new and unexpected in light of these previous results (see Response 2.2).
While related, these two studies considered particular cases that we do not consider at length. 
For example, Ubeda et al consider ESD-GSD transitions. 
Nevertheless, these transitions involve increased linkage with a haploid selected locus, which is precedent for our Conclusion 3A (which considers only GSD-GSD transitions).
We now mention this in the Introduction (L59-61) and again in the Discussion (L432).  
Similarly, Kozielska et al involves a special case of GSD transitions during which sex-linkage goes from perfect to none, which is is an extreme example of our Conclusions 2A \&2B.
This is twice mentioned (L64-67, L237). 
Despite this precedent, we find the permissive parameter space allowing GSD-GSD transitions surprising. 
%Especially because their main conclusion involved the strong influence of sex-ratio selection, which is explicitly not a driver during many such transitions. 
In addition, our Conclusions 4A and 4B reconcile Kozielska et al and Ubeda et al, explaining that sex-ratio selection is not an overwhelming force driving transitions in sex determination. 
We believe that our new presentation makes the novel and unexpected features of our results much more apparent. 
%To be clear, we do not claim to be the first to include both haploid and diploid selection to a model investigating the evolution of new sex-chromosomes and we do not consider the addition of sex-specific selection or analytical results for previously numerical models to be our most significant contributions (again, see Responses 2.1 and 2.2 and below for more detail on what these are). 
}

@@

\noindent\subsubsection{}
The findings in [SOO] claimed to be surprising (``Surprisingly, we find that neither force (sex ratio selection nor associations with genes that have sex-specific effects) dominates the spread of new sex-determining systems alone.") ... [but this result is rather similar to those in Ubeda et al 2015].

\noindent\paragraph{Response}
\textcolor{blue}{
Here we were referring to the result that $\lambda_{W'}^{XY} = \lambda_{Y'}^{ZW}$ with haploid selection, even when sex-ratio biases evolve to become either stronger or weaker (e.g., $\lambda_{Y'}^{ZW}$ vs. $\lambda_{W'}^{XY}$ with $R<r$ and male meiotic drive) or when sex-ratio biases are either present or absent (e.g., $\lambda_{W'}^{XY}$ vs. $\lambda_{Y'}^{ZW}$ with $r<1/2$, $R=1/2$, and male meiotic drive).  
Step 2 in Ubeda et al 2015 shows that associations with a haploid selected locus can favour a new sex-determining allele despite causing biased sex ratios. 
Nevertheless, it was difficult for us to intuit whether the presence/absence of ancestral sex-ratio bias would have a negligible or overwhelming effect on the evolution of new sex-determining alleles. 
For example, step 3 of Ubeda et al involves the spread of an autosomal meiotic drive suppressor favoured by Fisherian sex-ratio selection and Kozielska et al 2010 show that ancestral sex-ratio bias can favour new sex determining systems that balance the sex ratio.
These previous results are now discussed in the Introduction (L57-69). 
We found our result, that these two forces are exactly equal (with selection weak relative to recombination), such that there is no difference between the propensity for XY-ZW or ZW-XY transitions, to be surprising.
To clarify our meaning we have rephrased (in Author summary): ``Surprisingly, we find the two forces (selection to equalize the sex-ratio and the benefits of hitch-hiking with driven alleles that distort the sex ratio) will often be equally strong, and thus neither is sufficient to explain the spread of new sex-determining systems in every case." 
We believe that the re-framing of the Introduction/Discussion and the emphasis given to specific Conclusions will help to clarify this point throughout the paper.
}

### 2018-05-01 LaTeX/main_revised.tex
rearranged conclusions once again

From the arguments above we conclude that
\vspace{0.5cm}

@@

\nameref{app:inv_cond}) shows that Conclusion 2A holds with or without haploid selection, and further demonstrates that
\vspace{0.5cm}

\noindent\textit{Conclusion 2B:}
\textbf{
With tight linkage between a selected locus and the ancestral sex-determining locus ($r\approx0$), haploid selection can drive trans-GSD transitions (XY $\leftrightarrow$ ZW) that weaken sex-linkage ($r<R$), even when selection is sexually-antagonistic with directional selection in each diploid sex.
These trans-GSD transitions are driven by selection to equalize the sex ratio.
As in Conclusion 2A, such transitions potentially shift sex determination to a different chromosome ($R=1/2$) and can lead to polymorphic sex-determination systems.}
\textcolor{red}{[consistent with special case of Kozielska et al 2010]}

@@

%However, with haploid selection and some ancestral sex-linkage ($r<1/2$; allowing allele frequency differences on the X and Y), the additional term in Eq \eqref{eq:lambda_neoW} can be positive.
%This implies, 
%\vspace{0.5cm}
\noindent\textit{Conclusion 3B:}\
\textbf{New sex-determining alleles (causing trans-GSD transitions, XY $\leftrightarrow$ ZW) can spread even if they arise less closely linked with a locus that experiences selection than the ancestral-sex-determining locus is ($r<R$), requiring haploid selection and some initial sex-linkage ($r<1/2$) as necessary conditions.}
\textcolor{red}{[consistent with examples in Ubeda et al 2015]}  
\vspace{0.5cm}

@@

%Under Fisherian sex-ratio selection, autosomal modifiers favour equal investment in male and female offspring, i.e., a 1:1 sex ratio~\cite{Fisher:1930wy,Charnov:1982wg,West:2009we}. 
%A novel environmental sex-determiner that causes half of its carriers to become female and half to become male ($k=1/2$) will be in males half of the time and in females half of the time (like an autosome).
%In addition, these novel sex-determination alleles equalize the sex ratio and therefore one might expect them to be favoured by Fisherian sex-ratio selection when the resident sex ratio is biased.
%However, assuming weak selection, we find that the growth rate of a rare, dominant offspring-controlled neo-ESD allele that produces males or females with equal probability ($k=1/2$) is
Of particular interest are ESD mutations that cause half of their carriers to develop as females and half as males ($k=1/2$), creating equal sex ratios.
The spread of such mutations is determined by
\begin{equation}
\lambda_{ESD'}^{(XY)} =1+ \frac{1}{2}\frac{(\lambda_{Y'\rvert R=1/2}^{(XY)}-1) + (\lambda_{W'\rvert R=1/2}^{(XY)}-1)}{2} + O\left(\epsilon^3\right),
%V_A \frac{r}{1-2r}  \frac{s_{f} \left( s^\male - 3 s^\female\right) }{8 {\left( s_{f} + s_{m} \right)}^2} t_{m}^2.
\label{eq:lambda_ESD}
\end{equation}
\noindent
where $\lambda_{Y'\rvert R=1/2}^{(XY)}$ and $\lambda_{W'\rvert R=1/2}^{(XY)}$ represent $\lambda_{Y'}^{(XY)}$ and $\lambda_{W'}^{(XY)}$ when evaluated at $R=1/2$ (Equations \ref{eq:lambda_neoY} and \ref{eq:lambda_neoW}).
That is, recombination between the selected locus and the novel sex-determining locus, $R$, does not enter into the $k=1/2$ results.
This is because sex is essentially randomized each generation, %as if the novel sex-determiner were on an unlinked site that half the time acts as a neo-W and half the time as a neo-Y.
preventing associations from building up between alleles at locus $\mathbf{A}$ and sex. 
Eq \eqref{eq:lambda_ESD} shows that the ESD mutation gets half of the fitness of a feminizing mutation (neo-W) and half of the fitness of a masculinizing mutation (neo-Y), but only has an effect one half of the time (the other half of the time it produces the same sex as the ancestral system would have). 
As discussed above, $\lambda_{Y'\rvert R=1/2}^{(XY)}$ is necessarily less than one (cis-GSD transition), but $\lambda_{W'\rvert R=1/2}^{(XY)}$ (trans-GSD transition) can be greater than one if there is haploid selection (see Conclusion 2B).
That is, with haploid selection, an allele causing environmental-sex-determination can invade an ancestrally-XY system because it generates females that are either rare or have high fitness, in the same manner as a neo-W (likewise, ESD invades a ZW system for the same reasons a neo-Y can).
%An important result from Eq \eqref{eq:lambda_ESD} is that ESD can invade if there is haploid selection. 
%When evaluated at $R=1/2$, $\Lambda_{Y',XY}$ is necessarily less than one, but $\Lambda_{W',XY}$ can be greater than one if there is haploid selection, as discussed above. 
%Previous studies where ESD is favoured have typically assumed that environmental conditions (e.g., maternal condition, mate quality, age, or host size) can differentially affect the fitness of males versus females such that ESD invades because it allows sex determination to depend on the environment \citep[reviewed in][]{Charnov:1982wg,Bull:1983vi,West:2009we}. 
%Here, ESD mutations can spread because they generate females that are either rare or have high fitness, in the same manner as a neo-W.

@@

New sex-determination systems are typically expected to spread when they equalise the sex ratio and/or when they increase linkage with loci that experience sex-differences in selection~\cite{Blaser2012, vanDoorn2014re}.
In accordance with the latter mechanism, we find that sex-differences in selection at the haploid stage can favour trans-GSD transitions that tighten sex-linkage (Conclusion 2A). %, as exemplified in~\cite{Ubeda:2015fx} for ESD-GSD transitions. 
However, we also find that trans-GSD transitions can be favoured in cases where linkage with the sex-determining locus is actually weakened (Conclusions 1A-1C \& 2B, Fig \ref{fig:SexAntagTighter} \& \ref{fig:regionplots}), especially when there is haploid selection (Conclusion 2B, Figs \ref{fig:Combination_Centimorgans} \& \ref{fig:SexRatioBad}). 
Furthermore, we show that the spread of new sex-determination systems is not dominated by selection to balance the sex ratio (Conclusions 3 \& 4, Fig \ref{fig:SexRatioBad}).

%We find that the spread of new sex-determining systems cannot be simply predicted from their effect on the sex ratio. 
On one hand, sex-ratio biases caused by haploid selection can facilitate trans-GSD transitions or transitions from genetic to environmental sex determination~\cite{Kozielska:2010vm}. 
For instance, alleles favoured by haploid selection in males often become associated with the Y allele, which leads to an ancestral male-biased zygotic sex ratio.
This male bias increases the potential for a neo-W or ESD allele to invade (Table \ref{tab:haplotype_growth}), which can equalize the sex ratio (e.g., see Fig \ref{fig:SexRatioBad}B, for related examples see~\cite{Kozielska:2010vm}).
%We find that sex-ratio biases caused by haploid selection can also affect transitions between sex-determining systems (e.g., see $\zeta$ terms in Table \ref{tab:haplotype_growth}). 
%However, the neo-W may not equalize the sex ratio (e.g., if the driving allele also drives in females or affects competition among eggs)
On the other hand, sex-ratio selection can be overwhelmed by additional selective effects, preventing a neo-W or ESD allele from invading, even if it would balance the sex ratio (e.g., when selection also acts in opposite directions in male and female diploids, Table \ref{tab:specialcases}). %and equalizing the sex ratio.
Indeed, transitions between sex-determining systems can generate stronger sex-ratio biases (e.g., Fig \ref{fig:SexRatioBad}A and step 1 in~\cite{Ubeda:2015fx}).
%For example, where a neo-Y invades and is linked with a locus that experiences haploid selection in male gametes, the sex ratio evolves to become biased \citep[e.g., see Fig \ref{fig:SexRatioBad}A and step 1 in][]{Ubeda:2015fx}.
%For example, in an ancestral ZW system, haploid selection in males can allow a linked neo-Y to invade, despite the fact that it creates a male-biased sex ratio because the neo-Y becomes associated with the driven allele (Fig \ref{fig:SexRatioBad}A).
%\textcolor{blue}{and figure 1, where the end of one panel is the same as the beginning of another, however, this figure may change).}
%What we would like to stress is that sex-ratio selection alone cannot predict when new sex-determining systems can evolve.
Significantly, with weak selection, we find that there is no difference in conditions allowing XY to ZW and ZW to XY transitions (Conclusion 3) %, indicating that transitions in sex determination are not predominantly predicted by their effect on the sex ratio (i.e., the sex-ratio bias created by male haploid selection facilitates the spread of a neo-W allele into an XY system to the same degree that male haploid selection drives the spread of a neo-Y into a ZW system with a 1:1 sex ratio). 
%An asymmetry can develop when sex-linkage is tight (e.g., Fig \ref{fig:Combination_Centimorgans} near -25cM and 25cM) but under most circumstances we do not predict asymmetry between XY to ZW and ZW to XY transitions despite the presence/absence of sex-ratio selection. 
even when haploid selection always acts in the same sex (e.g., males).  
That is, the sex-ratio bias created by male haploid selection facilitates the spread of a neo-W allele into an XY system to the same degree that male haploid selection drives the spread of a neo-Y into a ZW system with a 1:1 sex ratio.
%Thus, haploid selection can favour trans-GSD transitions both via sex-ratio selection and via selection on alleles associated with the new sex-determining allele, and these selective pressures are often predicted to be of equal magnitude.

@@

New sex-determining alleles have previously been shown to spread when they arise in linkage with loci that experience sex differences in selection because beneficial associations build up between alleles that determine sex and alleles that are favoured in that sex~\cite{vanDoorn:2007eu,vanDoorn:2010hu,Ubeda:2015fx,Muralidhar2018}. 
In support of this hypothesis, researchers have identified genes on recently derived sex chromosomes that might be under sexually-antagonistic selection~\cite{Lindholm:2002dw,Tripathi:2009cw,Ser:2010iq}. 
However, we show that, if selected loci are tightly linked to the ancestral sex-determining locus, they can drive trans-GSD transitions which reduce sex-linkage (Conclusions 1A-1C \& 2B), thus widening the range of genomic locations that could be driving observed trans-GSD transitions.
In addition, we find that polymorphic sex determining systems (X, Y, and neo-W alleles all segregating) can be maintained when a selected locus is tightly linked to the ancestral sex-determining system (e.g., \nameref{fig:freqAll}B and \nameref{fig:freqAll}C), which is not possible with loose linkage~\cite{vanDoorn:2010hu}. 
These conclusions apply in cases with or without haploid selection.

@@

We have shown that tight sex-linkage and haploid selection can drive previously unexpected transitions between sex-determining systems.
In particular, both can select for new sex-determining loci that are more loosely linked to loci under selection (Conclusions 1A-1C \& 2B). 
In addition, haploid selection can cause transitions in GSD analogous to those caused by purely sexually-antagonistic selection, eliminating the need for differences in selection between male and female diploids (Conclusion 2A).
We conclude that haploid selection should be considered as a pivotal factor driving transitions between sex-determining systems. 
Further, transitions involving haploid selection can be driven by sex-ratio selection or cause sex-ratio biases to evolve; to leading order, selection to balance the sex ratio and the benefits of hitch-hiking with haploid selected alleles are of equal magnitude (Conclusions 3 \& 4). 
Overall, our results suggest several new scenarios under which new sex-determining systems are favoured, which could help to explain why the evolution of sex-determining systems is so dynamic.

>>>>

From the arguments above and in \nameref{app:inv_cond} we conclude that
\vspace{0.5cm}

@@

\nameref{app:inv_cond} shows that Conclusion 2A holds with or without haploid selection and further demonstrates that
\vspace{0.5cm}

\noindent\textit{Conclusion 2B:}
\textbf{
With tight linkage between a haploid selected locus and the ancestral sex-determining locus ($r\approx0$), trans-GSD transitions (XY $\leftrightarrow$ ZW) can be driven by selection to better specialize on the ancestrally-non-heterogametic sex and/or sex-ratio selection.
Such transitions reduce sex ratio bias.
}
\textcolor{red}{[consistent with special case of Kozielska et al 2010]}

@@

%However, with haploid selection and some ancestral sex-linkage ($r<1/2$; allowing allele frequency differences on the X and Y), the additional term in Eq \eqref{eq:lambda_neoW} can be positive.
%This implies, 
%\vspace{0.5cm}
\noindent\textit{Conclusion 3B:}\
\textbf{New sex-determining alleles (causing trans-GSD transitions, XY $\leftrightarrow$ ZW) can spread even if they arise less closely linked with a locus that experiences selection than the ancestral-sex-determining locus is ($r<R$), requiring haploid selection and some initial sex-linkage ($r<1/2$).}
\textcolor{red}{[consistent with examples in Ubeda et al 2015]}  
\vspace{0.5cm}

@@

%Under Fisherian sex-ratio selection, autosomal modifiers favour equal investment in male and female offspring, i.e., a 1:1 sex ratio~\cite{Fisher:1930wy,Charnov:1982wg,West:2009we}. 
%A novel environmental sex-determiner that causes half of its carriers to become female and half to become male ($k=1/2$) will be in males half of the time and in females half of the time (like an autosome).
%In addition, these novel sex-determination alleles equalize the sex ratio and therefore one might expect them to be favoured by Fisherian sex-ratio selection when the resident sex ratio is biased.
%However, assuming weak selection, we find that the growth rate of a rare, dominant offspring-controlled neo-ESD allele that produces males or females with equal probability ($k=1/2$) is
Of particular interest are ESD mutations that cause half of their carriers to develop as females and half as males ($k=1/2$), creating equal sex ratios.
The spread of such mutations is determined by
\begin{equation}
\lambda_{ESD'}^{(XY)} =1+ \frac{1}{2}\frac{(\lambda_{Y'\rvert R=1/2}^{(XY)}-1) + (\lambda_{W'\rvert R=1/2}^{(XY)}-1)}{2} + O\left(\epsilon^3\right),
%V_A \frac{r}{1-2r}  \frac{s_{f} \left( s^\male - 3 s^\female\right) }{8 {\left( s_{f} + s_{m} \right)}^2} t_{m}^2.
\label{eq:lambda_ESD}
\end{equation}
\noindent
where $\lambda_{Y'\rvert R=1/2}^{(XY)}$ and $\lambda_{W'\rvert R=1/2}^{(XY)}$ represent $\lambda_{Y'}^{(XY)}$ and $\lambda_{W'}^{(XY)}$ when evaluated at $R=1/2$ (Equations \ref{eq:lambda_neoY} and \ref{eq:lambda_neoW}).
That is, recombination between the selected locus and the novel sex-determining locus, $R$, does not enter into the $k=1/2$ results.
This is because sex is essentially randomized each generation, %as if the novel sex-determiner were on an unlinked site that half the time acts as a neo-W and half the time as a neo-Y.
preventing associations from building up between alleles at locus $\mathbf{A}$ and sex. 
Eq \eqref{eq:lambda_ESD} shows that the ESD mutation gets half of the fitness of a feminizing mutation (neo-W) and half of the fitness of a masculinizing mutation (neo-Y), but only has an effect one half of the time (the other half of the time it produces the same sex as the ancestral system would have). 
As discussed above, $\lambda_{Y'\rvert R=1/2}^{(XY)}$ is necessarily less than one (cis-GSD transition), but $\lambda_{W'\rvert R=1/2}^{(XY)}$ (trans-GSD transition) can be greater than one if there is haploid selection (see Conclusion 3B).
That is, with haploid selection, an allele causing environmental-sex-determination can invade an ancestrally-XY system because it generates females that are either rare or have high fitness, in the same manner as a neo-W (likewise, ESD invades a ZW system for the same reasons a neo-Y can).
%An important result from Eq \eqref{eq:lambda_ESD} is that ESD can invade if there is haploid selection. 
%When evaluated at $R=1/2$, $\Lambda_{Y',XY}$ is necessarily less than one, but $\Lambda_{W',XY}$ can be greater than one if there is haploid selection, as discussed above. 
%Previous studies where ESD is favoured have typically assumed that environmental conditions (e.g., maternal condition, mate quality, age, or host size) can differentially affect the fitness of males versus females such that ESD invades because it allows sex determination to depend on the environment \citep[reviewed in][]{Charnov:1982wg,Bull:1983vi,West:2009we}. 
%Here, ESD mutations can spread because they generate females that are either rare or have high fitness, in the same manner as a neo-W.

@@

New sex-determination systems are typically expected to spread when they equalise the sex ratio and/or when they increase linkage with loci that experience sex-differences in selection~\cite{Blaser2012, vanDoorn2014re}.
In accordance with the latter mechanism, we find that sex-differences in selection at the haploid stage can favour trans-GSD transitions that tighten sex-linkage (Conclusion 3A). %, as exemplified in~\cite{Ubeda:2015fx} for ESD-GSD transitions. 
However, we also find that trans-GSD transitions can be favoured in cases where linkage with the sex-determining locus is actually weakened (Conclusions 2A, 2B \& 3B, Fig \ref{fig:SexAntagTighter} \& \ref{fig:regionplots}), especially when there is haploid selection (Conclusion 3B, Figs \ref{fig:Combination_Centimorgans} \& \ref{fig:SexRatioBad}). 
Furthermore, we show that the spread of new sex-determination systems is not dominated by selection to balance the sex ratio (Conclusions 4A \& 4B, Fig \ref{fig:SexRatioBad}).

%We find that the spread of new sex-determining systems cannot be simply predicted from their effect on the sex ratio. 
On one hand, sex-ratio biases caused by haploid selection can facilitate trans-GSD transitions or transitions from genetic to environmental sex determination~\cite{Kozielska:2010vm}. 
For instance, alleles favoured by haploid selection in males often become associated with the Y allele, which leads to an ancestral male-biased zygotic sex ratio.
This male bias increases the potential for a neo-W or ESD allele to invade (Table \ref{tab:haplotype_growth}), which can equalize the sex ratio (e.g., see Fig \ref{fig:SexRatioBad}B, for related examples see~\cite{Kozielska:2010vm}).
%We find that sex-ratio biases caused by haploid selection can also affect transitions between sex-determining systems (e.g., see $\zeta$ terms in Table \ref{tab:haplotype_growth}). 
%However, the neo-W may not equalize the sex ratio (e.g., if the driving allele also drives in females or affects competition among eggs)
On the other hand, sex-ratio selection can be overwhelmed by additional selective effects, preventing a neo-W or ESD allele from invading, even if it would balance the sex ratio (e.g., when selection also acts in opposite directions in male and female diploids, Table \ref{tab:specialcases}). %and equalizing the sex ratio.
Indeed, transitions between sex-determining systems can generate stronger sex-ratio biases (e.g., Fig \ref{fig:SexRatioBad}A and step 1 in~\cite{Ubeda:2015fx}).
%For example, where a neo-Y invades and is linked with a locus that experiences haploid selection in male gametes, the sex ratio evolves to become biased \citep[e.g., see Fig \ref{fig:SexRatioBad}A and step 1 in][]{Ubeda:2015fx}.
%For example, in an ancestral ZW system, haploid selection in males can allow a linked neo-Y to invade, despite the fact that it creates a male-biased sex ratio because the neo-Y becomes associated with the driven allele (Fig \ref{fig:SexRatioBad}A).
%\textcolor{blue}{and figure 1, where the end of one panel is the same as the beginning of another, however, this figure may change).}
%What we would like to stress is that sex-ratio selection alone cannot predict when new sex-determining systems can evolve.
Significantly, with weak selection, we find that there is no difference in conditions allowing XY to ZW and ZW to XY transitions (Conclusion 4A) %, indicating that transitions in sex determination are not predominantly predicted by their effect on the sex ratio (i.e., the sex-ratio bias created by male haploid selection facilitates the spread of a neo-W allele into an XY system to the same degree that male haploid selection drives the spread of a neo-Y into a ZW system with a 1:1 sex ratio). 
%An asymmetry can develop when sex-linkage is tight (e.g., Fig \ref{fig:Combination_Centimorgans} near -25cM and 25cM) but under most circumstances we do not predict asymmetry between XY to ZW and ZW to XY transitions despite the presence/absence of sex-ratio selection. 
even when haploid selection always acts in the same sex (e.g., males).  
That is, the sex-ratio bias created by male haploid selection facilitates the spread of a neo-W allele into an XY system to the same degree that male haploid selection drives the spread of a neo-Y into a ZW system with a 1:1 sex ratio.
%Thus, haploid selection can favour trans-GSD transitions both via sex-ratio selection and via selection on alleles associated with the new sex-determining allele, and these selective pressures are often predicted to be of equal magnitude.

@@

New sex-determining alleles have previously been shown to spread when they arise in linkage with loci that experience sex differences in selection because beneficial associations build up between alleles that determine sex and alleles that are favoured in that sex~\cite{vanDoorn:2007eu,vanDoorn:2010hu,Ubeda:2015fx,Muralidhar2018}. 
In support of this hypothesis, researchers have identified genes on recently derived sex chromosomes that might be under sexually-antagonistic selection~\cite{Lindholm:2002dw,Tripathi:2009cw,Ser:2010iq}. 
However, we show that, if selected loci are tightly linked to the ancestral sex-determining locus, they can drive trans-GSD transitions which reduce sex-linkage (Conclusions 2A, 2B, \& 3B), thus widening the range of genomic locations that could be driving observed trans-GSD transitions.
In addition, we find that polymorphic sex determining systems (X, Y, and neo-W alleles all segregating) can be maintained when a selected locus is tightly linked to the ancestral sex-determining system (e.g., \nameref{fig:freqAll}B and \nameref{fig:freqAll}C), which is not possible with loose linkage~\cite{vanDoorn:2010hu}. 
These conclusions apply in cases with or without haploid selection.

@@

We have shown that tight sex-linkage and haploid selection can drive previously unexpected transitions between sex-determining systems.
In particular, both can select for new sex-determining loci that are more loosely linked to loci under selection (Conclusions 2A, 2B \& 3B). 
In addition, haploid selection can cause transitions in GSD analogous to those caused by purely sexually-antagonistic selection, eliminating the need for differences in selection between male and female diploids (Conclusion 2B \& 3B).
We conclude that haploid selection should be considered as a pivotal factor driving transitions between sex-determining systems. 
Further, transitions involving haploid selection can be driven by sex-ratio selection or cause sex-ratio biases to evolve; to leading order, selection to balance the sex ratio and the benefits of hitch-hiking with haploid selected alleles are of equal magnitude (Conclusions 4A \& 4B). 
Overall, our results suggest several new scenarios under which new sex-determining systems are favoured, which could help to explain why the evolution of sex-determining systems is so dynamic.
